\chapter*{Acknowledgments}
\addcontentsline{toc}{chapter}{Acknowledgments}

Undertaking and completing PhD is a journey of many people, and there are many people who deserve credit.
Particularly I wish to thank the efforts of my supervisors, my colleagues and also my family for making the completion of this PhD a possibility.

I would like to thank my supervisor and panel chair Sylvie Thi\'{e}baux, for her patience and kindness in taking regular time out of her schedule to talk and oversee my progress and discuss directions, and particularly for encouraging me to stick to a specific research direction.
I would like to thank Paul Scott, for his level-headedness and encouraging me to be tenacious about creating and producing quality research.
I would like to thank Archie Chapman, for sharing his enthusiasm about game theory and power systems and for providing me with creative research ideas and oversight on their development.

Throughout my research I was privileged to be a witness to the ARENA funded CONSORT project, to witness the implementation of grid-connected battery technology on Tasmania's Bruny island, and I wish to thank Sylvie, Paul and Archie, as well as the many other participants on the project for allowing me to witness such a trial of the implementation of new technology.
I also would like to thank Lachlan Blackhall for providing me with experience with power systems and opportunity for engineering experience through Reposit Power, which led to the opportunity for me to conduct this PhD.
I would also like to thank some of the friends and colleagues I have made at the ANU College of Engineering and Computer Science for supporting and encouraging me, particularly: Charles Gretton, Josh Milthorpe, Ahmad Attarha, Mahdi Noori, Masoume Mahmoodi.

I would like to thank my family, particularly my father Roderick and my brother Joseph for being with me throughout my years of rocky development.
And lastly I would like to express gratitude for my mother who passed away throughout the PhD, for showing me love and kindness.
