\chapter{Some background philosophy on Distributive Justice}
\label{cha:background}\label{sec:philosophy}

%\begin{displayquote}
%My claim is merely that there is no single fundamental principle that determines or provides guidance on what justice requires in relation to the distribution of access to overall advantage.\cite{mason2006levelling} 
%\end{displayquote}

In considering the question of what electricity market structure \textit{should} be implemented, it is essential to atleast acknowledge that there exists a wide range of moral and practical factors that bear on the question.

In this chapter we briefly consider some of the moral considerations that frame the question, however in doing so we must make clear that the deliberate brevity of this chapter is not to suggest that these moral considerations are not important and worthy of much greater treatment. But that we are not philosophers, and hence we intend only to acknowledge the moral situation with some of its depth.

%However these considerations and concepts are crucial and range much wider than the specific context of electricity markets, though unfortunately we cannot give a comprehensive treatment of them here.

The moral and ethical side of our question is associated with a branch of moral philosophy called \textit{Distributive Justice}, which seeks to ask and make headway on the question of how different kinds of resources (such as money/power/goods/etc) should be distributed in society; and this is the a broader question.

We begin by acknowledging the ambiguous nature of moral knowledge, and the approach that we take notwithstanding, before addressing some different categories of moral intuitions surrounding the nature of distributive justice.
Particularly we give brief surveys about ideas of: formal equality, social equality, efficiency, and proportionality.

\subsection{A philosophical prelude}

We begin by noting there is a long history of philosophical scepticism about the nature of moral knowledge and judgements, and Distributive Justice is not exceptional in this regard.

An example historical argument is Hume's `Guillotine' \cite{HumeGutenberg}\footnote{``For as this ought, or ought not, expresses some new relation or affirmation, it is necessary that it should be observed and explained; and at the same time that a reason should be given, for what seems altogether inconceivable, how this new relation can be a deduction from others, which are entirely different from it. But as authors do not commonly use this precaution, I shall presume to recommend it to the readers; and am persuaded, that this small attention would subvert all the vulgar systems of morality, and let us see, that the distinction of vice and virtue is not founded merely on the relations of objects, nor is perceived by reason.'' T3.1.1} which is often read as stating that: no material facts about how the physical world \textit{is}, by-themself, could ever seem to logically imply any claim about how the world (or its material components) \textit{should} be.

Another historic argument is G.E. Moore's open-question argument \cite{MooreGutenberg}\footnote{
``Moreover any one can easily convince himself by inspection that the predicate of this proposition - `good' - is positively different from the notion of `desiring to desire' which enters into its subject: `That we should desire to desire A is good' is not merely equivalent to `That A should be good is good.' ... clearly that we have two different notions before our minds.''Ch1:13\\
``If I am asked `What is good?' my answer is that good is good, and that is the end of the matter. Or if I am asked `How is good to be defined?' my answer is that it cannot be defined, and that is all I have to say about it.'' Ch 1:6}, which argues that for anything which defines what is morally good, then a question-about or statement-of that equivalence would only be tautology.

Such arguments are probably best used as discussion-starters today, however, talking about the ontological nature and the basis of moral knowledge is not our focus.
%Whether or not moral truth ultimately amounts to collective sentiment, or reduces to  statements of prudence, or is metaphysically identical/grounded in some deeper objective fact is beyond the scope of this work.
Instead, our focus tends towards discussion around the moral views that people are likely to have upon reflection; and we give an extremely brief survey some of the attitudes expressed in literature.

What is quite evident, is that different people have different conceptions of how the world should be, and not all of these conceptions are compatible with each other.
That any particular ethical system is likely to be rooted in a specific focus (as encoded by principles, maxims, cultural narrative, language etc) and will yield outcomes that may be disagreeable to some people and agreeable to others.
%In constructing any particular answer to the question of what counts as a good/better/fair mechanism we will necessarily isolate those who would disagree.
%And for this we must make a modest apology.

%We could just directly constructing an answer to the question \textit{de novo}, but such an approach would miss a lot of humility and leave open the question why those elements should be selected. And so the situation leads us to give at least some consideration the space of peoples moral intuitions before settling down upon a judgment.

We attempt to give a brief survey to address what we believe are often some of the elements that feature in people's moral thinking, and attempt to develop a novel synthesis about electricity systems, which bears some relevance to these moral considerations.

While we must acknowledge the moral ambiguity inherent in the question of electricity allocation, we contend that this does not mean that any answer is simply \textit{as good} as any other. But only that we believe that the suitability of our answer is not something we can totally demonstrate, in principle.

Let us begin.

\subsection{Introducing moral factors about distributive justice}

The choice of centralised market structures and processes can be seen as a choice between methods of allocating resources between multiple parties in a system based on the parties interaction within it.
In this context the choice of the market structures and also the resultant likely distribution of resources can be viewed as being morally/socially desirable or undesirable based on a number of factors.
What constitutes a desirable distribution of resources?

%In the process of discussing a choice of system it is important to come to a relatively clear understanding of what concepts are in play. And in these sections we will attempt to briefly survey and break-down some of the surrounding moral concepts.

Throughout time there have been an array of philosophers who have discussed ideas surrounding the moral distribution of resources and capital; and one of the major ideas surrounding the ethics of distribution is \textit{Equality}.

\subsection{Moral equality}

\begin{displayquote}
``A common characteristic of virtually all the approaches to the ethics of social arrangements that have stood the test of time is to want equality of \textit{something}... They are all `egalitarians' in some essential way ... To see the battle as one between thoes `in favor of' and thoes `against' equality (as the problem is often posed in the literature) is to miss something central to the subject."\cite[Chapter 1]{18084} 
\end{displayquote}

\begin{displayquote}
``for all men have some natural inclination to justice ... what is equal appears just, and is so; but not to all; only among those who are equals: and what is unequal appears just, and is so; but not to all, only amongst those who are unequals;\\
which circumstance some people neglect, and therefore judge ill; the reason for which is, they judge for themselves, and every one almost is the worst judge in his own cause." \cite[Politics, chapter III.9]{AristotleGutenberg}
\end{displayquote}


People tend to believe that they are, should be, or be treated, `equal' in some sense.
And this broad conception has changed throughout time and place in history \cite{themeaningofequalitycapaldi}.
From at least as far back as Aristotle \cite{AristotleGutenberg}\footnote{see section quote}, notions and concepts about equality have come from across culture and peoples, and between Spiritual \footnote{across multiple religions, eg. in Islam ``No Arab is superior to a non-Arab, no colored person to a  white person, or a white person to a colored person except by Taqwa (piety)." [Ahmad and At-Tirmithi], and in Christianity, St Paul's Galatians 3:28 ``There is neither Jew nor Gentile, neither slave nor free, nor is there male and female, for you are all one in Christ Jesus'' (NIV) } and the Materialist \footnote{Such as in Engel's Anti-D\"{u}hring Part 1 Chapter 10 ``The idea that all men, as men, have something in common, and that to that extent they are equal, is of course primeval. But the modern demand for equality is something entirely different from that; this consists rather in deducing from that common quality of being human, from that equality of men as men, a claim to equal political social status for all human beings''} thought.
Throughout the ages the way in which equality in society has been constructed and implemented has varied dramatically.% - even the abolition of slavery is recent by comparison.

There is something appealing about the idea of Equality between people.
From the asthetic perspective equality is an ideal with a simple structure. From a humanitarian perspective equality is associated with relief from envy and want. From the social perspective it is associated with community and solidarity.% From the philosophical perspective equality is directly associated with fundamental and core ideas.

%What is morality and justice? and what Rights and Freedoms does it require? and are people actually equal in any morally relevant way?

On a practical level, the divergences between people's ideas of equality can be seen as regarding what things should-be equal (when, where and for whom); and also what should be done about inequalities as they may exist.

The question: ``Equality of what?'' can have many answers, some of which are commonly held and seldom controversial today, such as might be gleamed from the United Nation's declaration of Human Rights: Equality before the Law, Democratic Equality to vote, equal freedoms to marry and to live, etc.
However more controversial answers tend to have broader social and political scope, such as: Equality of Opportunity, and Equality of Welfare and/or Economic Equality.% And the controversial nature of these ideas notably come to the fore in discussions such as surrounding affirmative action initiatives and also political socialism.

For some, equality is a contestable notion, or an ideal for the direction of efforts in narrow and specific contexts, but for others equality is an attainable and far-reaching goal with multifaceted implications across social spheres \cite{walzer2008spheres,millerandwalzer,baker1992arguing}.

There are many ways in which equality of specific things have been argued; particularly, a person can advocate for an equality in a given context \textit{directly} or alternatively as a means to some other end, ie. \textit{instrumentally}.
%For instance, as a practical measure in some settings, and also as an extension of more fundamental moral principles.
So, What amounts to an equal allocation can also instrumentally satisfy other values, for instance, David Miller \cite{equalityandjustice:1998} %argues for the practicality of giving equal remuneration between hypothetical employees in the context of uncertainty about how much each of them deserved (and/or lack of means about distinguishing them). And also 
gives some broad examples of reasons for equalities in society: for aesthetic and pro-social reasons, because it can be a sufficiently practical and simple social contract, or because it might be politically inevitable etc.
%These (and others) may count as instrumental or prudent reasons to value the implementation of an equality in a specific context, but in other contexts these same reasons could potentially point to other arrangements.
Alternatively, there have been arguments \textit{directly} for the equality of specific things, particularly from (or in light of) more abstract concepts such the notion that people have \textit{equal moral worth} \cite{doallpersonshaveequalmoralworth} or \textit{moral equality}.

Though it is difficult to define\footnote{for instance, Discussion about \textit{who} has equal moral worth (or alternatively \textit{how/why} they do) seems to occasionally to turn into a discussion about the moral rights of animals, \cite{doallpersonshaveequalmoralworth}}, the notion that people have equal moral worth is sometimes seen not to logically imply any very specific kind of equality of measure.
\begin{displayquote}
The distinction between ``equal treatment'' and ``treatment as equals'' expresses this difference between offering people the same treatment, and acting in accordance with the fact that they are moral equals. Equal status does not constrain us to a set of identical actions regardless of our differences.\cite{whatisbasicequalitynathan}
\end{displayquote}

The question about what things should be equal (rights, freedoms, duties, responsibilities etc, and for whom and when) can be seen as forming a large component of the various moral systems.
It is also sometimes seen that moral equality simply cannot be a logical premise for these questions.

\begin{displayquote}
The idea of moral equality, while fundamental, is too abstract to serve as a premise from which we deduce a theory of justice. What we have in political argument is not a single premise and then competing deductions, but rather a single concept and then competing conceptions or interpretations of it. Each theory of justice is not \textit{deduced from} the ideal of equality, but rather \textit{aspires to} it, and each theory can be judged by how well it succeeds in that aspiration.\cite{kymlicka2002contemporary}
\end{displayquote}

Thus it is perhaps better to take a more descriptive process to analysing equality, rather than a deductive one.
So for instance, expecting any particular person to give a precisely defined answer to the question ``Equality of what?'' may be asking too much; as even the phrases which people use in everyday life are seldom given exact specifications\footnote{Degrees of vagueness are well witnessed in everyday sentences, ``There are a gathering of people near that tree'', such as argued in the classic Sorties paradox\cite{frances_2018}}, let alone concepts pertaining to the spectra of possible societies.
In this way, the space of various contemporary political philosophies which faithfully attempt to construct and interpret some reasonable form of equality between persons has been described as belonging to an `egalitarian plateau'.\footnote{The phrase is originally attributed to Dworkin and subsequently adopted by others.}\cite{Brown2007}
Or conversely, while a specific ethical equality may not be agreed upon, perhaps there may be a more broadly accepted notion of what an `inequality' looks like,
particularly as it is sometimes blurred with the concept of a `social injustice'.\footnote{There is some debate as to when/where/how an inequality also becomes an injustice. It is possible to believe that an inequality constitutes an injustice directly, or perhaps that an inequality is proof (or perhaps only potential evidence) of a injustice in procedure or treatment. see parfit's concept of Telic vs. Deontic Egalitarianism.\cite{equalityandpriorityparfit}}

And in this way, the concept of Equality is seen as a vague notion that can be inclusive-of and also contrasted-against other views; such as those that emphasise the priority of resources to the poor, or such as emphasise alleviation of insufficiency among the poor; broadly termed ``prioritarianism'' and ``sufficientarianism'' respectively.\cite{sep-egalitarianism}\footnote{for good measure we might also consider Rawl's Theory of Justice \cite{rawls2005theory} as a specific kind of (layered) priority principle.}

Although the concept of equality is the subject of wider analysis, we will focus on two specific interpretations of equality which we feel can be made relevant to mechanisms for electricity allocation.

\subsection{Formal equality}\label{sec:formal_equality}

One of the most common interpretation of equality is that people should be subject to systems that treat them in a manner that is \textit{impartial}. The minimal idea is that an impartial system should not afford arbitrary or unjustified special treatment toward any particular individual/s. Hence that systems should operate by rules which are blind to particular identity and sensitive only to morally relevant characteristics.

Straightforward examples of this doctrine are embedded in anti-discrimination legislation, Mason \cite{mason2006levelling} gives some examples.  A person's religion for instance, is not generally a morally relevant characteristic for employability decisions, except in some example cases, such as in the performance of religious positions (eg. a Priest), or where a person's religious practice (such as wearing religious clothing, eg. a turban) may directly affect performance (such as on the construction site, requiring a hard-hat).
Many more examples exist, but what is notable is that where and when morally relevant characteristics arise is not always easy.

At a more general level, the idea of moral impartiality has been clarified by various thought experiments and also stated with moral maxims.
Particularly famous devices include Rawl's ``Original Position'', Kant's categorical imperatives, or various positions defined by hypothetical ideal sympathy and/or perfect detachment \cite{smithGutenberg, nla.cat-vn197822,10.2307/2103988}.
Additionally the idea of impartiality is perhaps somewhat mathematically expressible, in that people who are (in all the relevant ways) equal should be treated equally; and this is known as \textit{formal equality} \cite{whatisbasicequalitynathan}.

Although formal equality is occasionally seen as being an important part of a fair system, it is also sometimes seen to be insufficient to capture broader notions of equality and justice.

\begin{displayquote}
In its majestic equality, the law forbids rich and poor alike to sleep under bridges, beg in the streets and steal loaves of bread.\\
--Anatole France, Le Lys Rouge [The Red Lily] (1894), ch. 7
\end{displayquote}

Indeed, by imagination many kinds treatment or processes could be rationalised as being issued by impartial principles which are universally applied. Furthermore, not all kinds of desirable impartiality are mutually compatible, or perfectly achievable in practice.\cite{Hutchinson_2019}

Notwithstanding, formal equality can be seen as a basic doctrine that ascribes value to the incorporating degrees (and/or kinds) of impartiality into the design of social processes from the outset.%; even if they never truly reach perfect impartiality in practice.
% that social systems should be designed with an aim to treat people identically, but-for morally relevent characteristics.
%And thus there is perhaps some value in incorporating some degree (and/or kinds) of impartiality into the design of social processes from the outset; even if they never truly reach perfect impartiality in practice.
%in the fact that some degree (and/or kinds) of impartiality might be incorporated into the design of social processes from the outset; even if they never reach perfect impartiality in practice.
%It might also be profitable to consider the inequalities between the different kinds of rules (and treatement) that people can be effectively subject to \textit{in practice}, or effectively \textit{by consequence}.

In all developments in this thesis, the principle of formal equality - that individuals are treated equally but-for specific factors - is assumed.

\subsection{Equalities of social freedoms}

\begin{displayquote}
I want to emphasise what is, on my view, the most important object of egalitarian distribution, and that is \textit{power}. Of course power is not something which can be parcelled up and shared out like a commodity but we can properly talk of `the distribution of power' and this is, more than anything, the determinant of whether a community is authentically cooperative.\cite{TheSocialBasisofEquality:1998}
\end{displayquote}

There are different and interrelated ways of how to conceive of wider social equality, and one historic way of framing social equality is in terms of power.
For some people, the ideal of equality encodes the hope of a society free of abusive power relations that perpetuate social injustices.

One of more historically notable instances of this framing is featured in Marxist thought, which focuses on abusive economic power relations between social classes. This frame also shows up historically in feminist thought (eg. see \cite{Cudd2006-CUDAO}) where the inequalities of power between men \& women are considered as a form of oppressive dominance \& submission.\footnote{eg. MacKinnon writes ``difference is the velvet glove on the iron fist of domination. The problem is not that differences are not valued; the problem is that they are defined by power''\cite{mackinnon1989toward}.}

But what is notable is that neither Marxist nor feminist writers always viewed power itself, negatively.
For instance, Marx opposed private property (as capitalistic ownership) but seems to have had a more complex attitude toward property relations generally.\footnote{"the theory of the Communists may be summed up in the single sentence: Abolition of private property ... Do you mean the property of the petty artisan and of the small peasant, a form of property that preceded the bourgeois form? There is no need to abolish that"\cite{MarxGutenberg}\\"Property thus originally means no more than a human being's relation to his natural conditions of production as belonging to him, as his, as presupposed along with his own being; relations to them as natural presuppositions of his self, which only form, so to speak, his extended body."\cite[Notebook V]{marx1993grundrisse}}
Additionally some feminists occasionally consider power in the positive (or potentially neutral) language of \textit{empowerment} \cite{doi:10.1111/j.1527-2001.1998.tb01350.x}.

One of the features of power that is associated with abuse, is the exercise of `power over' other people, or `power to' do things which impinge apon other's rights.\cite{doi:10.1111/j.1527-2001.1998.tb01350.x}
But dissecting when and where an exercise of power constitutes an abusive or morally objectionable act may not be easy.
Particularly the `power to' do something is straightforwardly an example of a freedom, and one well known dichotomy exists between \textit{positive} and \textit{negative} freedoms\footnote{While vague, a negative freedom is associated with an absence of external obstacles to conducting the specific action, and a positive freedom is associated with the possibility (or actuality) of doing the act in accordance with one's will and purposes. The positive/negative dichotomy is also associated with what is or is not effortfull.\cite{Mossel2009-MOSNA}} particularly in the discussion of doing or allowing harm.
But even more broadly, freedoms can be considered as triadic relationships: a freedom \textit{of} a person, \textit{from} particular preventing conditions, \textit{to} do certain things.\cite{Negative_and_Positive_Freedom}

However different freedoms are not equally valued (or compatible), and some are esteemed by individuals and societies more than others.
Some would place an importance on political freedoms (to openly discuss, vote, and run for office) or economic freedoms (to work, to buy, sell and lease property), etc. %\footnote{indeed perhaps too many things, such as, as ownership of resources and money, occupying a social status, opportunity for advantage can be considered as being freedoms}
But particularly, the having and actualising of freedoms associated with the meeting of needs; such as basic needs (of shelter, food, etc - as at the bottom of Maslow's hierarchy) as well as higher needs (such as social belonging and self-actualisation); can be considered as defining of human wellbeing, and perhaps even a constituent of the state of having `Freedom' - the moral and political ideal.

Unfortunately most of these (and other) wider conceptions of societal Equality are beyond the scope of what we can earnestly engineer directly. But what we can do is to reflect and evaluate the influence that any proposed system might have on the freedoms of individuals and the wellbeing of society - and this is a task we attempt in later section \ref{sec:GNK_value_discussion}.

\subsection{Efficiency and utility maximisation}\label{sec:philosophy_efficiency}

%\begin{displayquote}
%Essentially, Utilitarianism sees persons as locations of their respective utilities %- as sites at which such activities as desiring and having pleasure and pain take place.
% ... Persons do not count as individuals in this any more than individual petrol tanks do in the analysis of the national consumption of petrolium.
%\cite{}
%\end{displayquote}

\begin{displayquote}It is the greatest happiness of the greatest number that is the measure of right and wrong. \textit{A Fragment on Government}, Jeremy Bentham\cite{bentham2001fragment}
\end{displayquote} 

It is occasionally thought that what is morally good for society should have some relationship with what is good for the individuals of society; and there is a question about how to characterise that relationship.
Historically what is morally good for individuals has been associated with such things as happiness \cite{burns2005happiness} or subjective welfare \cite{10.2307/2264894}, access to resources (such as electrical power) \cite{10.2307/2265047}, and/or opportunity for welfare \cite{10.2307/4320203}.%; particularly as these have been used as the object of various egalitarian and utilitarian formulations.

In more material contexts, what is good for an individual may be associated with: access to sufficient food and medicine, monetary yeilds, opportunities for educational attainment, probability of survival etc.\footnote{many thought experiments invoke lifeboat/class-room/triage/trolley-problem circumstances, where what is good for specific individuals is unambiguous}

But for whatever measures are considered to be relevant, the question of how these quantities should combine to bear on the broader moral judgement about what is good for society, has a variety of answers.
It is useful to illustrate the question by introducing the concept of utility as a quantification of what is good for individuals.

%What is morally good for individuals has defined the historical concept of `utility' and moral philosophies that focus on the utility of individuals are generally called `utilitarian'.
The concept of utility has changed over time\footnote{historically and notably held by famous utilitarians such as John Stuart Mill\cite{MillGutenberg} and Jeremy Bentham (see section quote), who defined it in terms of happiness or pleasure/pain \cite{bentham1823introduction}}
but minimally it is concieved as a measure of the strength of the preference (or value) that a specific person does (or alternatively should, rationally, pragmatically and/or morally) attach to different possible outcomes.% - in a specific context or perhaps more generally.
\footnote{the concept can be seen to extend from the consideration that such preferences should be transitive and comparable between people. If a person prefers A to B, and also B to C, then they ought to also prefer A to C. And if one person can prefer A `more' than another prefers B - then it remains a task of invention to associate numbers to the strength of these preferences over outcomes.}
For this definition, the sum of utility is straightforwardly one of many examples of a \textit{collective utility function}, a function that aggregates the utility of individuals and hence is a possible target for moral decision making.\cite{TheoriesofValueAggregation}

As another example, some people's egalitarian intuitions might be satisfied by moral decision making that affords equal utility for all individuals - and in some contexts this may be appropriate, even though it might not maximise the sum of utility. This outcome could be constructed as an alternative collective utility function.

What is to be realised is that maximising one collective utility function does not necessarily maximise the other.
Particularly this is made clear in the famous 'levelling-down objection', which is the objection that, a person implementing a strict Egalitarian distribution of utility would potentially prefer a world in which every single person had less, if it were more equal. \cite{temkin_2003, equalityandpriorityparfit}.

These considerations frame some possible articulations of the broader contrast between the values of Equality and 'Efficiency'; where the specific concept of efficiency considered in the levelling-down objection is \textit{Pareto optimality}.
Specifically, an outcome is defined as being pareto optimal if there does not exist another outcome which is better than it for every person\footnote{more specifically that is, strictly better for atleast one, better or equal for the rest.}.
Pareto optimality is one commonly discussed and formalised efficiency condition, and is a property which we satisfy in our subsequent developments.
Particularly as it is the case that maximising the sum of utility is pareto optimal \cite{TheoriesofValueAggregation} and forms an axiom in our treatment (as it is given by our efficiency axiom \eqref{myeq} in chapter \ref{cha:new_solution}).

We must note that there is a range of respectable ways to aggregate the moral preferences of people which we cannot address here.
Additionally, in various realistic situations the difference between sensible outcomes which are better or best for more people, and what is more equal for them, can be the subject of dispute (such as in the medical field\cite{Reidpath2012,RePEc:chy:respap:120cherp}, the provision of welfare \cite{10.2307/27522452}, and economics \cite{10.1093/oep/gpz040}). Notwithstanding the philosophical distinction between the various articulations of efficiency between people and equality among them, can remain.

%We note however that there is a range of possible collective utility functions, particularly important is the space of prioritarian solutions, where which by weighted functions prefers utility imputation to thoes worst off, which in the extreme case prioritises only the worst off, which is an articulation of Rawl's difference principle and for which some regard as being 'the' egalitarian solution.

%Indeed these approaches are hardly exhaustive, and the result of the considerations of which one is appropriate may be highly context dependant. (people should have food of sufficient nutrition, public housing should be allocated as priority to thoes in most need of it. government subsidies should maximise the total welfare of the nation etc.) 

\subsection{Fairness by proportionality to some reference point}\label{sec:reference_points}

There are various ways in which what is considered fair is determined in relation to what is normative, and/or related to the various counterfactual events that could happen.
%These may be related to the implementation of moral rules, related to what is a natural and expected condition, or moral or culturally constructed expected of people, and may in turn relate to diferences in situations which people could effect.
One of the more basic and famous examples of this relationship is given by Mill's harm principle:

\begin{displayquote}
That the only purpose for which power can be rightfully exercised over any member of a civilised community, against his will, is to prevent harm to others. - John Stuart Mill, On Liberty, chapter 1 \cite{Mill2Gutenberg}
\end{displayquote}

%Can may be considered by many as a condition of fairness and justice.
Insofar as the harm principle is accepted, there subsequently remains a question about defining where and when harm occurs.
Among its features, harm is often considered to be negative and defined with respect to a more normative (and potentially counterfactual) `unharmed' state.
However in more complicated cases, it is not always clear what the more normative `unharmed' state should be.%, and sometimes thus who is harmed by whom.

But also reversely, there are also various conceptions of justice which involve compensation for providing benefit to others.
For instance, in Business ethics there is a viewpoint where an employee's just wage should be in proportion to their contribution to the value and productivity of the firm.\cite{sternberg2000just}
In this context, the contribution may by measured by profit relative to their absence and/or by the replacement cost of contracting equivalent work (potentially depending on which ever is more pertinent).
Particularly, the idea that people should be rewarded in proportion to their contribution to the social good above their absence individually, is most directly rendered by the Vickrey-Clarke-Groves (VCG) mechanism - see section \ref{sec:solutions_VCG}.

We note that this relation to a reference points may be somewhat associated with the ideas of deservedness, compensation, reward and/or proportionality; however these relation is not necessarily exact.
For instance, the idea of compensation is most often associated with damages or injury to others, but less commonly associated detracting gain that occurs to them, the notion of reward is often associated specifically with incentivising particular pro-social behaviours among individuals, and to some extent the same is true of the validity of deservedness claims.

Furthermore in considering the notion of proportionality, the one primary question which occurs, is `proportionality with respect to what?' - ie. what should the reference point be?

However, thinking in terms of proportionality may be slightly misleading as example of a normative reference point is also found in Business ethics.
Particularly there is a viewpoint where a business transaction (such as pay for an employee's work) is considered justified if it was attained by a process of truly-free negotiation between the parties, such as to make all parties better off than they would be otherwise.\cite{ExecutiveCompensationUnjustorJustRight} 
This truly-free exchange (sometimes called \textit{euvolentary exchange}) is particularly defined by the fact that all parties could have realistically elected to walk-away from the negotiation.\cite{Guzman2019}
The event that would be triggered if the parties did not successfully negotiate is sometimes called the `disagreement event', the existance of which is the normative reference point that determines the morality of the transaction.
Some of these ideas are mathematically rendered by various bargaining solution concepts, such as Nash bargaining - see section \ref{sec:solutions_bargaining}.

In these cases we can see instances where the morality of an event is defined by normative reference points. And this dynamic can extend even to groups of individuals, as we might consider ways that groups of individuals might be exploited even if their individual interactions are truly-free.
\begin{displayquote}
``This is the main aim of John Roemer's work on [Marxist] exploitation.\\... %He defines Marxist exploitation, not in terms of surplus transfer, but in terms of unequal access to the means of production. Whether one is exploited or not, on his view, depends on whether one would be better off in a hypothetical situation of distributive equality -- namely, where one withdraw with one's labour and per capita share of external resources. 
If we view the different groups in the economy as players in a game whose rules are defined by existing property-relations then a group is exploited if its members would do better if they stopped playing the game, and withdrew their per capita share of external resources and started playing their own game.''\cite{kymlicka2002contemporary}
\end{displayquote}
And this idea of allocation exceeding what any group could achieve if they withdrew to cooperate among themselves, is most directly articulated and formalised by the \textit{Core} solution concept of cooperative game theory - see section \ref{sec:cooperative_game_theory_part}.

Similarly, another solution concept in cooperative game theory is the \textit{Shapley Value} (see section \ref{subsec:the_shapley_value}), which can be summarised as allocating compensation in proportion to each individuals contribution (in expectation) to group welfare above their absence, under uncertainty about the presence of other group members.

The consideration that the relevant normative reference point occurs under uncertainty (or in expectation), is featured in other formulations.
For instance, if we expect that euvolentary market exchanges would normally occur at a certain market price, then we may consider that the moral trading of goods would occur at this market price.
One famous example of this idea is featured in John Locke's short essay Venditio \cite{locke2003locke} where it is argued that a fair price for something is simply its normal market price at its location. (see section \ref{sec:solutions_LMP} for elaboration)

%The idea that moral trading is defined by normalised market conditions then raises the question of how normalised market conditions should be determined,
%In economics there are several approaches to integrating the dynamics of market forces, but one of the most historic frameworks is economic marginalism, and most particularly Locational Marginal Pricing (LMP) - see section \ref{sec:solutions_LMP}.\\

%https://reconstructingeconomics.com/2014/06/06/venditio-by-john-locke/

In all these cases, the morality of a situation is defined with respect to various reference points, which of these reference points are most relevant (and to what extent) may depend on a range of practical and moral factors.

It is also worth noting that these decisions about normative reference points can be defined by social policies %(such as minimum wages and award conditions)
 and embodied in moral codes and standards; and in this thesis, we consider development of a novel synthesis that extends from a specific reference point -- see Chapter \ref{cha:new_solution}. 

%It is also worth noting that the decision about these reference points can be made clear by social policies such as minimum wages and award conditions, and these rules 
%particularly famous is Rule Utilitarianism.

%\subsection{Wider moral considerations}

%There is wide range of perspectives on distributive justice which we can only begin to survey in this chapter.
%And there are many other positions that we could address, including conceptions of fairness as issuing from envy-freeness, ideas of fairness defined by equal share of social surplus, and all the various ways these viewpoints can intersect.%, such as rule utilitarianism.

%However, amongst the plurality of the various moral formulations, there are some contextualising frames which can be considered, such as the role that moral justice actually plays in society, in terms of evolution and psychology.
%The function of justice can be considered in the context of evolution; such as the idea that our drive towards particular moral norms and principles might have extended from repeated play evolutionary equilibria in the past. \cite{Binmore10785}
%For instance, systems of justice might be seen as encoding the evolved social expectations of reciprocity, and hence surving a purpose of mitigating the biological drive towards violence and retribution.\cite{doi:10.1300/J135v02n04}
%Additionally there is a possible psychological function, where, moral attitudes about the world is potentially a part of people's psychology \cite{doi:10.1037/1089-2680.6.1.25, doi:10.1080/00981389.2019.1640337}
%It is also sometimes witnessed that what people morally think should/ought happen is associated with what they \textit{will} to be.\cite{doi:10.1080/13869795.2016.1212395, sep-moral-motivation, LITZ2009695}

\subsection{A summary on our philosophy of distributive justice}


There is wide range of perspectives on distributive justice which we can only begin to survey in this chapter.
And there are many other positions that we could address, including conceptions of fairness as issuing from envy-freeness, ideas of fairness defined by equal share of social surplus, and all the various ways these viewpoints can intersect.%, such as rule utilitarianism.

Moral considerations are at the heart of the question of how electricity and monetary payments should be distributed.
Unfortunately questions such as these dont have analytically demonstrable answers, but we can consider various flavours of moral ideas which people might assert.
Particularly we summarise some of the various conceptions of Equality, Formal Equality, and Equality concerned with social freedoms.
Additionally we frame some of the concepts of Efficiency particularly as it is contrast against Equality.
And highlight the ways in which morality can be defined in relation to various normative reference points.

We focus particularly on these factors as our further developments relate directly to them.
Our solution obeys formal equality principles, maximises efficiency, and is formalised by reference to normative reference points defined by idealised competition -- see chapter \ref{cha:new_solution}.

%Summary what you discussed in this chapter, and mention the story in next
%chapter. Readers should roughly understand what your thesis takes about by only reading
%words at the beginning and the end (Summary) of each chapter.



